======================================

#title: "IndividualAtivityAnalysis.Rmd"
#author: "M Dugda"
#date: "Saturday, June 04, 2016"
#output: html_document
===================================

Creating Dirctory, downloading and unzipping data
======================================
```{r echo=TRUE} 
Url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip" 
if (!file.exists("data")) dir.create("data")
if (!file.exists("data/data.zip")) download.file(url, destfile = "data/data.zip", mode="wb")
if (!file.exists("data/activity.csv")) unzip("data/data.zip", exdir="data")
```

#Load the data and check its structure and summary
```{r echo =TRUE}
activity <- read.csv("data/activity.csv",quote="\"")
str(activity)
head(activity)
summary(activity)
activity$date <- as.Date(activity$date)
activity_ign <- subset(activity, !is.na(activity$steps))
```

#What is mean total number of steps taken per day?
```{r}
dailysum <- tapply(activity_ign$steps, activity_ign$date, sum, na.rm=TRUE, simplify=T)
dailysum <- dailysum[!is.na(dailysum)]

hist(x=dailysum,
     col="blue",
     breaks=20,
     xlab="Daily total steps",
     ylab="Frequency",
     main="The distribution of daily total")
mean(dailysum)
median(dailysum)
```

#What is the average daily activity pattern?
```{r}
int_avg <- tapply(activity_ign$steps, activity_ign$interval, mean, na.rm=TRUE, simplify=T)
activity_ia <- data.frame(interval=as.integer(names(int_avg)), avg=int_avg)

with(activity_ia,
     plot(interval,
          avg,
          type="l",
          xlab="5-minute intervals",
          ylab="average steps in the interval across all days"))
```

#Imputing missing values
```{r echo=TRUE} 
activity_impute <- activity
ndx <- is.na(activity_impute$steps)
int_avg <- tapply(activity_ign$steps, activity_ign$interval, mean, na.rm=TRUE, simplify=T)
activity_impute$steps[ndx] <- int_avg[as.character(activity_impute$interval[ndx])]

new_dailysum <- tapply(activity_impute$steps, activity_impute$date, sum, na.rm=TRUE, simplify=T)

hist(x=new_dailysum,
     col="green",
     breaks=20,
     xlab="daily steps",
     ylab="frequency",
     main="The distribution of daily total (with missing data imputed)")
     
mean(new_dailysum)

median(new_dailysum)
```

Are there differences in activity patterns between weekdays and weekends?

# helper function to decide if a day is a week day or not
```{r check if day is a weekday, echo=TRUE} 
is_weekday <- function(d) {
    wd <- weekdays(d)
    ifelse (wd == "Saturday" | wd == "Sunday", "weekend", "weekday")
}

wx <- sapply(activity_impute$date, is_weekday)
activity_impute$wk <- as.factor(wx)
head(activity_impute)
```

#make a panel plot containing a time series plot

```{r echo=TRUE}
wk_activity <- aggregate(steps ~ wk+interval, data=activity_impute, FUN=mean)
library(lattice)
xyplot(steps ~ interval | factor(wk),
       layout = c(1, 2),
       xlab="Interval",
       ylab="Number of steps",
       type="l",
       lty=1,
       data=wk_activity)
```

# The END of the Rmd Code